---
title: "Testing Set Features Visualization"
output: html_notebook
---

This notebook examines the relationship between different features of the data for distinguishing viral from nonviral sequences.

Please reach out to James Riddell (riddell.26@buckeyemail.osu.edu) or
Bridget Hegarty (beh53@case.edu) regarding any issues, or open an issue on github.

```{r setup-library}
library(ggplot2)
library(plyr)
library(reshape2)
library(viridis)
library(tidyr)
library(dplyr)
library(readr)
library(data.table)
```

```{r}
viruses <- read_tsv("../IntermediaryFiles/viral_tools_combined.tsv")
```

```{r}
colnames(viruses)
```

```{r}
ggplot(viruses, aes(x=RNA, y=viral)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("RNA VS2 Score") +
  ylab("VS2 Viral Score")

ggplot(viruses, aes(x=RNA, y=viral)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~confusion_matrix_high_MCC, scales = "free") +
  xlab("RNA VS2 Score") +
  ylab("VS2 Viral Score")

ggplot(viruses, aes(x=NCLDV, y=viral)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("NCLDV VS2 Score") +
  ylab("VS2 Viral Score")

ggplot(viruses, aes(x=NCLDV, y=viral)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~confusion_matrix_high_MCC, scales = "free") +
  xlab("NCLDV VS2 Score") +
  ylab("VS2 Viral Score")

ggplot(viruses, aes(x=lavidaviridae, y=viral)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Lavidaviridae VS2 Score") +
  ylab("VS2 Viral Score")

ggplot(viruses, aes(x=lavidaviridae, y=viral)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~confusion_matrix_high_MCC, scales = "free") +
  xlab("Lavidaviridae VS2 Score") +
  ylab("VS2 Viral Score")
```

# Important features by sequence type

```{r}
pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
```

```{r}
ggplot(viruses, aes(x=hallmark, y=checkv_viral_genes)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Number of Hallmark Genes") +
  ylab("Number of Viral Genes")
```

```{r}
ggplot(viruses, aes(x=checkv_host_genes, y=checkv_viral_genes)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Number of Host Genes") +
  ylab("Number of Viral Genes")
```

```{r}
ggplot(viruses, aes(x=percent_unknown, y=percent_viral)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Percentage of Genes Unknown") +
  ylab("Percentage of Genes Viral")
```

```{r}
ggplot(viruses, aes(x=percent_unknown, y=checkv_length)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Percentage of Genes Unknown") +
  ylab("Length of Sequence")
```

```{r}
ggplot(viruses, aes(x=percent_viral, y=checkv_length)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Percentage of Genes Viral") +
  ylab("Length of Sequence") 
```

```{r}
ggplot(viruses, aes(x=hallmark, y=checkv_length)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Number of Hallmark Genes") +
  ylab("Length of Sequence") 
```

```{r}
ggplot(viruses, aes(x=checkv_host_genes, y=checkv_length)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Length of Sequence") +
  ylab("Number of Host Genes")
```
```{r}
table(viruses$checkv_host_genes>=50, viruses$seqtype)
```



```{r}
ggplot(viruses, aes(x=checkv_length, y=checkv_completeness)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Length") +
  ylab("Completeness") 
```

```{r}
ggplot(viruses, aes(x=hallmark, y=checkv_completeness)) +
  geom_hex(bins = 30) +
  scale_fill_continuous(type = "viridis", trans="log10") +
  theme_bw() +
  facet_wrap(~seqtype, scales = "free") +
  xlab("Hallmark Genes") +
  ylab("Completeness") 
```

```{r}
table(viruses$seqtype[viruses$checkv_length>50000 & viruses$hallmark==0])/table(viruses$seqtype)
table(viruses$seqtype[((viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes) & viruses$checkv_provirus=="No"])/table(viruses$seqtype)
table(viruses$seqtype[viruses$checkv_viral_genes==0 & viruses$checkv_host_genes>=1])/table(viruses$seqtype)

table(viruses$seqtype[viruses$percent_viral>=50])/table(viruses$seqtype)
table(viruses$seqtype[viruses$percent_unknown>=75])/table(viruses$seqtype)
table(viruses$seqtype[viruses$percent_unknown>=75 & viruses$checkv_length<50000])/table(viruses$seqtype)
table(viruses$seqtype[viruses$hallmark>2])/table(viruses$seqtype[viruses$seqtype %in% unique(viruses$seqtype[viruses$hallmark>2])])
```

```{r}
table(viruses$seqtype, viruses$Kaiju_Viral)
```

```{r}
seqdata <- data.frame(seqtype=viruses$seqtype[!duplicated(viruses$contig)])

rownames(seqdata) <- viruses$contig[!duplicated(viruses$contig)]
```

```{r}
library(phyloseq)
```


```{r}
features_table <- viruses[viruses$Index==1,]
features_table <- features_table[,colnames(features_table) %in% c( 
                                              "checkv_viral_genes",
                                              "checkv_host_genes",
                                              "checkv_unknown_genes",
                                              "checkv_length",
                                              "checkv_completeness",
                                              "checkv_total_genes",
                                              "percent_host",
                                              "percent_viral",
                                              "hallmark",
                                              "percent_unknown"
                                              )]

features_table[is.na(features_table)] <- 0
ft_colnames <- colnames(features_table)
features_table <- t(features_table)
rownames(features_table) <- ft_colnames
colnames(features_table) <- rownames(seqdata)

physeq_pooled <- phyloseq(otu_table(features_table, taxa_are_rows = T))
```

```{r}
ordination <- phyloseq::ordinate(physeq =physeq_pooled, method = "PCoA", distance = "bray")
phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
                          shape="numtools", color="num_viruses") + 
  geom_point(size = 3) +
  theme_bw() +
  geom_label(label=seqdata$toolcombo)

phyloseq::plot_ordination(physeq = physeq_pooled, ordination = ordination,
                          shape="numtools", color="num_viruses") + 
  geom_point(size = 3) +
  theme_bw()
```

# Viral Addition Tuning Rules

```{r}
viruses_sankey_tv <- data.frame(seqtype=viruses$seqtype,
                             kj_cel=rep(0,nrow(viruses)),
                             hall=rep(0,nrow(viruses)),
                             pv=rep(0,nrow(viruses)),
                             cvl_pu=rep(0, nrow(viruses)))
```

```{r}
viruses_sankey_tv$kj_cel[viruses$Kaiju_Viral=="Viruses"] <- 0.5
viruses_sankey_tv$hall[viruses$hallmark>2] <- 0.5
viruses_sankey_tv$pv[viruses$percent_viral>=50] <- 0.5
viruses_sankey_tv$cvl_pu[viruses$checkv_length>50000 & viruses$percent_unknown<=75] <- 0.5    

viruses_sankey_tv$all <- rowSums(viruses_sankey_tv[,2:5])
```

```{r}
viruses_sankey_tv %>%
  count(seqtype, all) %>% spread(key = all, value=n)
```

```{r}
viruses_sankey_tv <- viruses_sankey_tv %>%
  count(seqtype, kj_cel, hall, pv, cvl_pu, all) %>%
  mutate(viral_score=factor(all))
```

```{r}
ggplot(viruses_sankey_tv,
       aes(axis1 = kj_cel, axis2 = hall, axis3 = pv, axis4 = cvl_pu, y=n)) +
  geom_alluvium(aes(fill=viral_score),
                width = 0, knot.pos = 0, reverse = FALSE) +
  geom_stratum(width = 1/5) +
  theme_bw() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)),
            reverse = FALSE) +
  theme(
        axis.text.x=element_text(size=14, angle = 90)
        ) +
  scale_x_continuous(breaks=c(1,2,3,4),
    labels=c("kaiju", "hallmark", "% viral", "% unknown")) +
  facet_wrap(~seqtype, scales="free_y") 
``` 

# Viral Removal Tuning Rules

```{r}
viruses_sankey_tnv <- data.frame(seqtype=viruses$seqtype,
                             kj_cel=rep(0,nrow(viruses)),
                             hg_pro=rep(0,nrow(viruses)),
                             vg_hg=rep(0,nrow(viruses)),
                             vg_hg_pro=rep(0, nrow(viruses)),
                             cvl_hm=rep(0, nrow(viruses)),
                             cvl_cp=rep(0,nrow(viruses)))
```

```{r}
viruses_sankey_tnv$kj_cel[viruses$Kaiju_Viral=="cellular organisms"] <- -0.5
viruses_sankey_tnv$hg_pro[viruses$checkv_host_genes>50 & viruses$provirus==FALSE] <- -1
viruses_sankey_tnv$vg_hg[viruses$checkv_viral_genes==0 & viruses$checkv_host_genes>=1] <- -1
viruses_sankey_tnv$vg_hg_pro[((viruses$checkv_viral_genes*3) <= viruses$checkv_host_genes) & viruses$provirus==FALSE] <- -1
viruses_sankey_tnv$cvl_hm[viruses$checkv_length>500000 & viruses$hallmark<=1] <- -1
viruses_sankey_tnv$cvl_cp[viruses$checkv_length>5000 & viruses$checkv_completeness<=75] <- -0.5             
viruses_sankey_tnv$all <- rowSums(viruses_sankey_tnv[,2:7])
```

```{r}
viruses_sankey_tnv %>%
  count(seqtype, all) %>% spread(key = all, value=n)
```


```{r}
viruses_sankey_tnv <- viruses_sankey_tnv %>%
  count(seqtype, kj_cel, hg_pro, vg_hg, vg_hg_pro, cvl_hm, cvl_cp, all) %>%
  mutate(viral_score=factor(all))
```

```{r}
ggplot(viruses_sankey_tnv,
       aes(axis1 = kj_cel, axis2 = hg_pro, axis3 = vg_hg, axis4 = cvl_hm, axis5=cvl_hm, axis6=cvl_cp, 
           y=n)) +
  geom_alluvium(aes(fill=viral_score),
                width = 0, knot.pos = 0, reverse = FALSE) +
  geom_stratum(width = 1/5) +
  theme_bw() +
  geom_text(stat = "stratum", aes(label = after_stat(stratum)),
            reverse = FALSE) +
  theme(
        axis.text.x=element_text(size=14, angle = 90)
        ) +
  scale_x_continuous(breaks=c(1,2,3,4,5,6),
    labels=c("kaiju", "host genes", "viral and host", "provirus",
             "hallmark", "complete")) +
  facet_wrap(~seqtype, scales="free_y") 
``` 













