---
title: "identify_viruses"
author: "James Riddell"
date: "8/25/2023"
output: html_document
---

This Rmarkdown file takes outputs from viral identification tools and merges
them into one file, "{dataset}_merged_viral_id_outputs.csv", which is the input
for identifying_viral_metrics_environmental.Rmd.

Please reach out to James Riddell (riddell.26@buckeyemail.osu.edu) or
Bridget Hegarty (beh53@case.edu) regarding any issues, or open an issue on github.

```{r setup-library}
library(ggplot2)
library(stringr)
library(plyr)
library(reshape2)
library(viridis)
library(tidyr)
library(dplyr)
library(readr)
library(data.table)

```

```{r}
# dataset name for file organization and outputs for metrics file
dataset_name <- "drinking_water"
#name from building_env_data_viral_features_set.Rmd
viruses_filename <- 'merged_viral_feature_outputs.csv'
# path of merged dataset
path = "/Users/riddellj/Documents/GitHub/VSTE/RunFilesForUsers/"
setwd(path)

# adjust to read file in.
viruses <- read_csv(paste(dataset_name, viruses_filename, sep='_'))
```

This section defines a viralness score "keep_score" based on the tool classifications. 
A final keep_score above 1 indicates we will keep that sequence and call it viral.

Rule
  subrule
  ...
  subrule

VIBRANT
    Quality == "Complete Circular": +1
    Quality == "High Quality Draft": +1
    Quality == "Medium Quality Draft": +1
    Quality == "Low Quality Draft" & provirus == TRUE: +0.5

Virsorter
    category ==  1 or 4: +1
    category == 2, 3, 5, or 6: +0.5
    
Virsorter2
    max_score >= 0.5: +0.5
    max_score >= 0.95: +0.5 (additive to the previous subrule)

DeepVirFinder:
    Score >= 0.7 & p-value <= 0.05 & checkv_length < 20kb: +0.5
    Score >= 0.9 & p-value <= 0.05 & checkv_length < 20kb: +0.5 (additive to the previous subrule)

Tuning Addition:
    Kaiju_viral == "Viruses" & kaiju_match_ratio >= 0.3: +1
    If %unknown >= 75 and checkv_length < 50000: +0.5
    If checkv_%viral >= 50 or vs2_%viral >= 50%: +0.5
    Hallmark >= 3: +1

Tuning - Not Viral:
    viral_genes == 0 & host_genes >= 1: -1
    If host_genes >50 and checkv_provirus == False: -1 
    If 3*viral_genes <= host_genes and  checkv_provirus == False: -1
    If length > 50,000 and hallmark <=1: -1


This script produces visualizations of these combined viral scorings and
includes ecological metrics like alpha diversity.

Users can decide which combination is appropriate for them and only need to run
the chunks for the appropriate tools.

Function that allows for the comparing pieces of the pipeline
```{r compute_viral_score}

#' The function compute_viral_score takes in a dataframe of sequences with viral
#' identification tool outputs and assigns a viralness score, "keep_score", based
#' on the classifications of each viral identification tool. The function returns
#' a list of keep_scores, one per observation of the input_seqs dataframe.

#' This function is used by assigning another column of the input_seqs dataframe
#' to the output of the function, associating a keep_score to each sequence.

#' Sequences with a keep_score >= 1 are considered viral. Refer to the manuscript
#' Figure 7 for which tool combinations to use based on the sample type. Refer
#' to the manuscript Figures 3-6 for performance of each combination.

#' parameters:
#'  input_seqs: dataframe with input sequence names and viral ID tool outputs
#'              from "build_features_table.Rmd"
#'  include_*:  boolean variables if the specified rule is to be used in
#'              computing the keep_score.
#'  tv_*:       sub-rules of tuning addition that can be individually turned on 
#'              or off.
#'  tnv_*:      sub-rules of tuning removal that can be individually turned on 
#'              or off.
compute_viral_score <- function(input_seqs,
                                include_vibrant=FALSE, 
                                include_virsorter=FALSE,
                                include_virsorter2=FALSE,
                                include_deepvirfinder=FALSE,
                                include_tuning_viral=FALSE,
                                tv_1=T, tv_2=T, tv_3=T, tv_4=T,
                                include_tuning_not_viral=FALSE,
                                ntv_1=T, ntv_2=T, ntv_3=T, ntv_4=T) {
  
  keep_score <- rep(0, nrow(input_seqs))
  if (include_vibrant) {
    keep_score[input_seqs$vibrant_quality=="high quality draft"] <- keep_score[input_seqs$vibrant_quality=="high quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="medium quality draft"] <- keep_score[input_seqs$vibrant_quality=="medium quality draft"] + 1
    keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$vibrant_prophage] <- keep_score[input_seqs$vibrant_quality=="low quality draft" & input_seqs$vibrant_prophage] + 0.5
  }
  
  if (include_virsorter2) {
    keep_score[input_seqs$max_score>=0.50] <- keep_score[input_seqs$max_score>=0.50] + 0.5
    keep_score[input_seqs$max_score>=0.95] <- keep_score[input_seqs$max_score>=0.95] + 0.5  
  }
  
  if (include_virsorter) {
    keep_score[input_seqs$category==1] <- keep_score[input_seqs$category==1] + 1
    keep_score[input_seqs$category==2] <- keep_score[input_seqs$category==2] + 0.5
    keep_score[input_seqs$category==3] <- keep_score[input_seqs$category==3] + 0.5
    keep_score[input_seqs$category==4] <- keep_score[input_seqs$category==4] + 1
    keep_score[input_seqs$category==5] <- keep_score[input_seqs$category==5] + 0.5 
    keep_score[input_seqs$category==6] <- keep_score[input_seqs$category==6] + 0.5 
  }
  
  if (include_deepvirfinder) {
    keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[(input_seqs$score>=0.9 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] + 0.5
    keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] <- keep_score[(input_seqs$score>=0.7 & input_seqs$pvalue<=0.05) & input_seqs$checkv_length<20000] + 0.5
  }
  
  if (include_tuning_viral) {
    keep_score <- tuning_viral(current_score=keep_score,
                               ins=input_seqs,
                         include_tv_1=tv_1,
                         include_tv_2=tv_2,
                         include_tv_3=tv_3,
                         include_tv_4=tv_4)
  }
  
  if (include_tuning_not_viral) {
    # tuning removal 
    keep_score <- tuning_not_viral(current_score=keep_score,
                               ins=input_seqs,
                         include_tnv_1=ntv_1,
                         include_tnv_2=ntv_2,
                         include_tnv_3=ntv_3,
                         include_tnv_4=ntv_4)
   
  }
  
  return(keep_score)
  
}
tuning_viral <- function(current_score, ins, 
                         include_tv_1=T,
                         include_tv_2=T,
                         include_tv_3=T,
                         include_tv_4=T) {
    # tuning addition
  if (include_tv_1) {
    current_score[ins$hallmark>=3] <- current_score[ins$hallmark>=3] + 1
  }
  if (include_tv_2) {
    current_score[ins$Kaiju_Viral=="Viruses" & ins$kaiju_match_ratio>=0.3] <- current_score[ins$Kaiju_Viral=="Viruses" & ins$kaiju_match_ratio>=0.3] + 1
  }
  if (include_tv_3) {
    current_score[ins$percent_unknown>=75 & ins$checkv_length<50000] <- current_score[ins$percent_unknown>=75 & ins$checkv_length<50000] + 0.5
  }
  if (include_tv_4) {
    current_score[ins$viral>=50 | ins$percent_viral>=50] <- current_score[ins$viral>=50 | ins$percent_viral>=50] + 0.5 
  }
  return(current_score)
}
tuning_not_viral <- function(current_score, ins, 
                         include_tnv_1=T,
                         include_tnv_2=T,
                         include_tnv_3=T,
                         include_tnv_4=T) {
    # tuning removal
  if (include_tnv_1) {
    current_score[(ins$checkv_host_genes>50) & !ins$checkv_provirus] <- -3
  }

  if (include_tnv_2) {
    current_score[ins$checkv_viral_genes==0 & ins$checkv_host_genes>=1] <- current_score[ins$checkv_viral_genes==0 & ins$checkv_host_genes>=1] - 1
  }
  if (include_tnv_3) {
    current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$checkv_provirus] <- current_score[((ins$checkv_viral_genes*3) <= ins$checkv_host_genes) & !ins$checkv_provirus] - 1
  }
  
  if (include_tnv_4) {
    current_score[ins$checkv_length>500000 & ins$hallmark<=1] <- -3
  }
  
  return(current_score)
}
```

The next code chunk is where you can create your own combination of tools,
or use one of the presets provided. Simply change each include_* rule to True
to include the rule, and False to exclude it.

```{r}
viruses$keep_score_all <- compute_viral_score(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = T,
                                              include_virsorter = T)
viruses$keep_score_high_recall <- compute_viral_score(viruses, include_deepvirfinder = T,
                                              include_vibrant = T,
                                              include_virsorter2 = T,
                                              include_tuning_viral = T,
                                              include_tuning_not_viral = F,
                                              include_virsorter = T)
viruses$keep_score_high_MCC <- compute_viral_score(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)
viruses$keep_score_high_precision <- compute_viral_score(viruses, include_deepvirfinder = F,
                                              include_vibrant = F,
                                              include_virsorter2 = T,
                                              include_tuning_viral = F,
                                              include_tuning_not_viral = T,
                                              include_virsorter = F)
```

```{r}
sum(viruses$keep_score_all>=1)
sum(viruses$keep_score_high_recall>=1)
sum(viruses$keep_score_high_precision>=1)
sum(viruses$keep_score_high_MCC>=1)
```

Write sequence names and viral scores to a file.
```{r}
write_csv(viruses[,c('contig','checkv_provirus','keep_score_all', 'keep_score_high_MCC', 'keep_score_high_precision', 'keep_score_high_recall')], paste(dataset_name, 'top_viral_scores.csv', sep="_"))
```

Write a list of sequences that are considered virus to file. Change high MCC
to whichever combination you choose to filter by. 

```{r}
virus_contig_list <- viruses[viruses$keep_score_high_MCC >= 1, c('contig', 'keep_score_high_MCC')]

write_csv(virus_contig_list, paste(dataset_name, 'virus_contig_list.csv', sep='_'))
```

Most will be done with using this tool at this stage. You can take the virus
contig list and use it to filter your fasta file for only contigs classified as
viral by whichever tool combination chosen above. 


Below are code chunks for comparing the number of viruses identified by different tool combinations.
These are not necessary to run, but beware that number of contigs called virus DOES NOT correlate to
total number of viruses, and may just be including more false positives. It is always recommended
to check contigs by looking at functional annotations (from DRAMv, for example) to confirm if 
the sequence is viral or not.


# Comparing all of the tool methods

```{r}
combos_list <- data.frame(toolcombo=rep(0, 64),
                          tune_not_viral=rep(0, 64),
                          DVF=rep(0, 64),
                          tune_viral=rep(0, 64),
                          VIBRANT=rep(0, 64),
                          VS=rep(0, 64),
                          VS2=rep(0, 64))
p <- 1
for (i in c(0,1)){
  for (j in c(0,1)){
    for (k in c(0,1)){
      for (l in c(0,1)){
        for (m in c(0,1)){
          for (n in c(0,1)){
            combos_list$toolcombo[p] <- paste(i,j,k,l,m,n)
            combos_list$toolcombo2[p] <- paste(if(i){"tv"}else{"0"},if(j){"DVF"}else{"0"},
                                               if(k){"tnv"}else{"0"},if(l){"VB"}else{"0"},
                                               if(m){"VS"}else{"0"},if(n){"VS2"}else{"0"})
            combos_list$tune_not_viral[p] <- k
            combos_list$DVF[p] <- j
            combos_list$tune_viral[p] <- i
            combos_list$VIBRANT[p] <- l
            combos_list$VS[p] <- m
            combos_list$VS2[p] <- n
            p <- p+1
          }
        }
      }
    }
  }
}
combos_list <- combos_list[-1,]
```

```{r}
viral_scores <- matrix(data=0, nrow=nrow(viruses), ncol=nrow(combos_list))
num_viruses <- data.frame(toolcombo=rep(0, nrow(combos_list)),
                          num_viruses=rep(0, nrow(combos_list)))
for (i in 1:nrow(combos_list)) {
  viral_scores[,i] <- compute_viral_score(viruses, include_vibrant = combos_list$VIBRANT[i],
                                            include_virsorter = combos_list$VS[i],
                                            include_virsorter2 = combos_list$VS2[i],
                                            include_tuning_viral = combos_list$tune_viral[i],
                                            include_tuning_not_viral = combos_list$tune_not_viral[i],
                                            include_deepvirfinder = combos_list$DVF[i])
  
  if (max(viral_scores[,i])<=0) {
    num_viruses$num_viruses[i] <- 0
  }
  else {
    num_viruses$num_viruses[i] <- table(viral_scores[,i]>=1)[[2]]
  }
  num_viruses$toolcombo[i] <- combos_list$toolcombo[i]
  
  num_viruses$toolcombo2[i] <- combos_list$toolcombo2[i]
}
num_viruses$numtools <- stringr::str_count(num_viruses$toolcombo, "1")
num_viruses <- num_viruses[order(num_viruses$num_viruses, decreasing=F),]
num_viruses$toolcombo <- factor(num_viruses$toolcombo, levels = unique(num_viruses$toolcombo))
num_viruses$toolcombo2 <- factor(num_viruses$toolcombo2, levels = unique(num_viruses$toolcombo2))
num_viruses$numtools <- as.factor(num_viruses$numtools)
```

This chunk will generate a plot of all tool combinations and how many viruses
were identified from the sample from each type.

```{r}
png(paste(dataset_name, 'virus_predicted.png', sep='_'))
pal <- ggthemes::tableau_color_pal(palette="Tableau 10", type="regular")
# add this
ggplot(num_viruses, aes(x=toolcombo, y=num_viruses, 
                                  color=numtools, fill=numtools)) +
  geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Num Viruses Predicted") +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(6)), 1))
dev.off()
```

```{r}
viral_scores <- matrix(data=0, nrow=nrow(viruses), ncol=nrow(combos_list))
num_assemblies <- dim(table(viruses$assembly))
assemblies <- names(table(viruses$assembly)) # add assemblies if not in viruses yet
num_viruses <- data.frame(toolcombo=rep(0, nrow(combos_list)*num_assemblies),
                          toolcombo2=rep(0, nrow(combos_list)*num_assemblies),
                          assembly=rep(0,nrow(combos_list)*num_assemblies),
                          num_viruses=rep(0, nrow(combos_list)*num_assemblies),
                          prop_vir=rep(0, nrow(combos_list)*num_assemblies)
                          )
n <- 1
for (i in 1:nrow(combos_list)) {
  viral_scores[,i] <- compute_viral_score(viruses, include_vibrant = combos_list$VIBRANT[i],
                                            include_virsorter = combos_list$VS[i],
                                            include_virsorter2 = combos_list$VS2[i],
                                            include_tuning_viral = combos_list$tune_viral[i],
                                            include_tuning_not_viral = combos_list$tune_not_viral[i],
                                            include_deepvirfinder = combos_list$DVF[i])
  for (j in 1:num_assemblies) {
    if (max(viral_scores[,i])<=0) {
      num_viruses$num_viruses[n] <- 0
    }
    else {
      num_viruses$num_viruses[n] <- sum(viral_scores[viruses$assembly==assemblies[j],i]>=1)
      num_viruses$prop_vir[n] <- num_viruses$num_viruses[n]/table(viruses$assembly)[[j]]
    }

    num_viruses$toolcombo[n] <- combos_list$toolcombo[i]
    num_viruses$toolcombo2[n] <- combos_list$toolcombo2[i]
    num_viruses$assembly[n] <- assemblies[j]

    n <- n+1
  }
}
num_viruses$numtools <- stringr::str_count(num_viruses$toolcombo, "1")
num_viruses <- num_viruses[order(num_viruses$prop_vir, decreasing=F),]
num_viruses$toolcombo <- factor(num_viruses$toolcombo, levels = unique(num_viruses$toolcombo))
num_viruses$toolcombo2 <- factor(num_viruses$toolcombo2, levels = unique(num_viruses$toolcombo2))
num_viruses$numtools <- as.factor(num_viruses$numtools)
```

```{r}
write_csv(num_viruses, paste(dataset_name, "num_viruses.csv", sep="_"))
```

```{r}
png(paste(dataset_name, 'prop_viral_assembly.png', sep='_'))
ggplot(num_viruses, aes(x=toolcombo, y=prop_vir, 
                                  shape=numtools, color=assembly, fill=assembly)) +
  geom_point() +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Proportion of Contigs >3kb Viral")
dev.off()
```

only plotting the average

```{r}
num_viruses$numtools <- as.numeric(num_viruses$numtools)
median_prop_vir_by_tool <- num_viruses %>% 
  select(toolcombo, prop_vir, numtools) %>%
  group_by(toolcombo) %>%
  summarize(median=median(prop_vir),
         #sd=sd(prop_vir),
         numtools=mean(numtools)) 
median_prop_vir_by_tool$toolcombo <- factor(median_prop_vir_by_tool$toolcombo, 
                                          levels = median_prop_vir_by_tool$toolcombo[order(median_prop_vir_by_tool$median, decreasing = F)])
median_prop_vir_by_tool$numtools <- as.factor(median_prop_vir_by_tool$numtools)
```

```{r}
png(paste(dataset_name, 'prop_viral_average.png', sep='_'))
ggplot(median_prop_vir_by_tool, aes(y=median, x=toolcombo,
                                  color=numtools, fill=numtools, shape=numtools))  +
  geom_hline(yintercept=0, color="grey") +
  geom_point() +
#  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
#                 position=position_dodge(.9)) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  xlab("Tool Combination (tv, DVF, tnv, VB, VS, VS2)") +
  ylab("Proportion of Contigs >3kb Viral") +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(6)), 1))
dev.off()
```
```{r}
png(paste(dataset_name, 'boxplot.png', sep='_'))
ggplot(median_prop_vir_by_tool, aes(x=numtools, y=median, 
                                  color=numtools, fill=numtools)) +
  geom_boxplot() +
  geom_point(alpha=0.5) +
  theme_light() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "bottom",
    axis.text.y=element_text(size=14),
    axis.text.x=element_text(size=14, angle = 90),
    legend.text=element_text(size=12),
    axis.title=element_text(size=16),
  ) +
  ylab("Proportion Viruses (%)") +
  xlab("Number of Tools") +
  scale_fill_manual(name="",
                     values = alpha(rev(pal(6)), 0.5)) +
  scale_color_manual(name="",
                     values = alpha(rev(pal(6)), 1))
dev.off()
```